### 2024-02-18 outputs/paper/paper.qmd
started paper and cleaning data

---
title: "Women and Gender Pay Gap "
subtitle: "Faulty World Bank Data Conflicts: A Canadian Example"
author: 
  - Sehar Bajwa
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/replicationpaper. "
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

SEE Discussion for paper idea. Basically there is world bank data about women and the labour market with data from 180+ countries that give a overview of the progression of women's outcomes in the labour market. However, when you compaire it with local Canadian data in the example of canada, the statistics don't really match up, therefore the data has a lot of probelms with collection. And we'd expand on that in our replication but the replication is not done (yet).

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

Can a woman choose where to live in the same way as a man?

Can a woman travel outside her home in the same way as a man?

Yes, they can but is a surface level question, because women do not travel the same way. Possible discussion of how transit designs are gender neutral, why designing for trip chaining is important.

---

[](https://static1.squarespace.com/static/57b25a1d579fb3a225546f2e/t/61f60fc732982f3789c5741d/1643521217247/finalreport.pdf)

---

Can a woman apply for a passport in the same way as a man?

---

Can a woman travel outside the country in the same way as a man?

---

**Workplace 100 (2010-2023)**

Can a woman get a job in the same way as a man?

[The gender wage gap in Canada: 1998 to 2018](https://www150.statcan.gc.ca/n1/pub/75-004-m/75-004-m2019004-eng.htm)

https://canadianwomen.org/blog/the-gender-pay-gap-is-a-myth/

Does the law prohibit discrimination in employment based on gender?

• In 2018, female employees aged 25 to 54 earned $4.13 (or 13.3%) less per hour, on average, than their male counterparts. In other words, these women earned $0.87 for every dollar earned by men.

- The two largest factors explaining the remaining gender wage gap in 2018 were the distribution of women and men across industries, and women’s overrepresentation in part-time work. These were also the largest explanatory factors behind the gap in 1998.
- Similar to other studies, nearly two-thirds of the gap in 2018 was unexplained. Possible explanations for this portion include gender differences in characteristics that were beyond the scope of this study, such as work experience, as well as unobservable factors, such as any gender-related biases.
- **Note: The difference between men’s and women’s average wages was found to be statistically significant at a 95% confidence level.**

**Is there legislation on sexual harassment in employment?**

he infographic "[Sexual misconduct and gender-based discrimination at work, 2020](https://www150.statcan.gc.ca/en/catalogue/11-627-M2021061)", released in 2021, indicated that 25% of women and 17% of men experienced sexual misconduct in their workplace in the previous year. Statistics Canada also reported that women were twice as likely as men to experience unwanted sexual behaviours in the workplace.

The largest proportion of women reported ever experiencing inappropriate sexualized behaviours in a workplace setting (44%), followed by discriminatory behaviours (20%) and sexual assault (13%).

Statistics Canada's [Centre for Gender, Diversity and Inclusion Statistics](https://www.statcan.gc.ca/en/topics-start/gender_diversity_and_inclusion), in collaboration with the Canadian Centre for Justice and Community Safety Statistics, has released a new data table on experiences of harassment and sexual assault in the workplace. This data table reports on the [Gender Results Framework](https://women-gender-equality.canada.ca/en/gender-results-framework.html) (GRF), introduced by the Government of Canada in 2018 to track the progress of gender equality in Canada. The new data table also includes data collected through the 2020 [Survey on Sexual Misconduct at Work](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5290) (SSMW).

**Are there criminal penalties or civil remedies for sexual harassment in employment?**

[The Daily — Gender Results Framework: A new data table on workplace harassment](https://www150.statcan.gc.ca/n1/daily-quotidien/240212/dq240212a-eng.htm)

https://www.learningtoendabuse.ca/research/national_survey_on_harassment_and_violence_at_work_in_canada/plain_text_documents/exploring_the_intersections_of_sexual_violence_and_precarious_work_a_literature_review.html

---

Pay (100)

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

\newpage

# References

### 2024-02-18 scripts/02-data_cleaning.R

@@ -1,44 +1,95 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans data for replication analysis (translated from Stata)
+# Author:  Marie Hyland, Simeon Djankov, and Pinelopi Koujianou Goldberg* (translated from the Stata by Luca Carnegie)
+# Date: 12 February 2023
+# Contact: luca.carnegie@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 01-download_data.R
 
 #### Workspace setup ####
 library(tidyverse)
+library(haven)
+library(readxl) 
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#### Read in data ####
+wb_data <- read_excel("inputs/data/repl_package/Data/WBL50YearPanelData.xlsx", sheet = "WBL1971-2020", range = "B2:AX9502", col_names = TRUE)
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
[…]

### 2024-02-18 outputs/paper/paper.qmd
cleaned data and edited paper!

@@
 library(tidyverse)
-library(palmerpenguins)
 ```
@@
 
-SEE Discussion for paper idea.
-Basically there is world bank data about women and the labour market with data from 180+ countries that give a overview of the progression of women's outcomes in the labour market.
-However, when you compaire it with local Canadian data in the example of canada, the statistics don't really match up, therefore the data has a lot of probelms with collection.
-And we'd expand on that in our replication but the replication is not done (yet).
-
 # Data {#sec-data}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+# Results {#sec-results}
 
@@
 
-## First discussion point {#sec-first-point}
+## First discussion point {#sec-point1}
 
-Can a woman choose where to live in the same way as a man?
+## Second discussion point {#sec-point2}
 
-Can a woman travel outside her home in the same way as a man?
-
-Yes, they can but is a surface level question, because women do not travel the same way. Possible discussion of how transit designs are gender neutral, why designing for trip chaining is important.
-
----
-
-[](https://static1.squarespace.com/static/57b25a1d579fb3a225546f2e/t/61f60fc732982f3789c5741d/1643521217247/finalreport.pdf)
-
----
-
-Can a woman apply for a passport in the same way as a man?
-
----
-
-Can a woman travel outside the country in the same way as a man?
-
----
-
-**Workplace 100 (2010-2023)**
-
-Can a woman get a job in the same way as a man?
-
-[The gender wage gap in Canada: 1998 to 2018](https://www150.statcan.gc.ca/n1/pub/75-004-m/75-004-m2019004-eng.htm)
-
-https://canadianwomen.org/blog/the-gender-pay-gap-is-a-myth/
-
-Does the law prohibit discrimination in employment based on gender?
-
-• In 2018, female employees aged 25 to 54 earned $4.13 (or 13.3%) less per hour, on average, than their male counterparts. In other words, these women earned $0.87 for every dollar earned by men.
-
-- The two largest factors explaining the remaining gender wage gap in 2018 were the distribution of women and men across industries, and women’s overrepresentation in part-time work.
-These were also the largest explanatory factors behind the gap in 1998.
-- Similar to other studies, nearly two-thirds of the gap in 2018 was unexplained.
-Possible explanations for this portion include gender differences in characteristics that were beyond the scope of this study, such as work experience, as well as unobservable factors, such as any gender-related biases.
-- **Note: The difference between men’s and women’s average wages was found to be statistically significant at a 95% confidence level.**
-
-**Is there legislation on sexual harassment in employment?**
-
-he infographic "[Sexual misconduct and gender-based discrimination at work, 2020](https://www150.statcan.gc.ca/en/catalogue/11-627-M2021061)", released in 2021, indicated that 25% of women and 17% of men experienced sexual misconduct in their workplace in the previous year.
-Statistics Canada also reported that women were twice as likely as men to experience unwanted sexual behaviours in the workplace.
-
-The largest proportion of women reported ever experiencing inappropriate sexualized behaviours in a workplace setting (44%), followed by discriminatory behaviours (20%) and sexual assault (13%).
-
-Statistics Canada's [Centre for Gender, Diversity and Inclusion Statistics](https://www.statcan.gc.ca/en/topics-start/gender_diversity_and_inclusion), in collaboration with the Canadian Centre for Justice and Community Safety Statistics, has released a new data table on experiences of harassment and sexual assault in the workplace.
-This data table reports on the [Gender Results Framework](https://women-gender-equality.canada.ca/en/gender-results-framework.html) (GRF), introduced by the Government of Canada in 2018 to track the progress of gender equality in Canada.
-The new data table also includes data collected through the 2020 [Survey on Sexual Misconduct at Work](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5290) (SSMW).
-
-**Are there criminal penalties or civil remedies for sexual harassment in employment?**
-
-[The Daily — Gender Results Framework: A new data table on workplace harassment](https://www150.statcan.gc.ca/n1/daily-quotidien/240212/dq240212a-eng.htm)
-
-https://www.learningtoendabuse.ca/research/national_survey_on_harassment_and_violence_at_work_in_canada/plain_text_documents/exploring_the_intersections_of_sexual_violence_and_precarious_work_a_literature_review.html
-
----
-
-Pay (100)
-
-## Second discussion point
-
-## Third discussion point
+## Third discussion point {#sec-point3}
 
 ## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 
@@
 
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
 \newpage

### 2024-02-18 scripts/00-simulate_data.R

@@ -20,11 +20,11 @@ set.seed(123)  # Setting seed for reproducibility
 report_years <- 2010:2020
 
 # Generating random scores for the specified columns 
-#Can only be 0, 25, 50, 75, 100 due to nature of survey
+# Note: Can only be 0, 25, 50, 75, 100 due to nature of survey
 scores <- sample(c(0, 25, 50, 75, 100), length(report_years), replace = TRUE)
 
 # Creating the data frame
-simulated_data <- data.frame(
+simulated_data <- tibble(
   Economy = rep("Canada", length(report_years)),
   Region = rep("High income: OECD", length(report_years)),
   Report_Year = report_years,
@@ -40,17 +40,18 @@ simulated_data <- data.frame(
 )
 
 # Testing simulated data
-#simulated_data$Report_Year |> min() == 2010
-#simulated_data$Report_Year |> max() == 2020
-#simulated_data$WBL_Index |> min() == 0
 
-#simulated_data$Mobility %in% c(0, 25, 50, 75, 100)
-#simulated_data$Workplace %in% c(0, 25, 50, 75, 100)
-#simulated_data$Pay %in% c(0, 25, 50, 75, 100)
-#simulated_data$Marriage %in% c(0, 25, 50, 75, 100)
-#simulated_data$Parenthood %in% c(0, 25, 50, 75, 100)
-#simulated_data$Entrepreneurship %in% c(0, 25, 50, 75, 100)
-#simulated_data$Assets %in% c(0, 25, 50, 75, 100)
-#simulated_data$Pension %in% c(0, 25, 50, 75, 100)
+simulated_data$Report_Year |> min() == 2010
+simulated_data$Report_Year |> max() == 2020
+simulated_data$WBL_Index |> min() == 0
+
+simulated_data$Mobility %in% c(0, 25, 50, 75, 100)
+simulated_data$Workplace %in% c(0, 25, 50, 75, 100)
+simulated_data$Pay %in% c(0, 25, 50, 75, 100)
+simulated_data$Marriage %in% c(0, 25, 50, 75, 100)
+simulated_data$Parenthood %in% c(0, 25, 50, 75, 100)
+simulated_data$Entrepreneurship %in% c(0, 25, 50, 75, 100)
+simulated_data$Assets %in% c(0, 25, 50, 75, 100)
+simulated_data$Pension %in% c(0, 25, 50, 75, 100)
 
 

### 2024-02-18 scripts/02-data_cleaning.R

@@ -11,8 +11,8 @@ library(tidyverse)
 library(haven)
 library(readxl) 
 
-#### Read in data ####
-wb_data <- read_excel("inputs/data/repl_package/Data/WBL50YearPanelData.xlsx", sheet = "WBL1971-2020", range = "B2:AX9502", col_names = TRUE)
+#### Read in data from Excel file ####
+wb_data <- read_excel("Replication Package/Data/WBL50YearPanelData.xlsx", sheet = "WBL1971-2020", range = "B2:AX9502", col_names = TRUE)
 
 # Rename clunky col names 
 wb_data <- wb_data |>
@@ -91,5 +91,24 @@ wb_data <- wb_data |> select(-all_of(yn))
 # Rename the new variables to the original names
 wb_data <- wb_data |> rename_with(~ str_remove(., "n"), starts_with("n"))
 
-# Clean up 
+# Clean up some labels. 
 
+wb_data <- wb_data |>
+            rename(`economy` = `Economy`) |>
+            rename(`countrycode` = `Code`)
+
+wb_data$countrycode <- ifelse(wb_data$economy == "West Bank and Gaza", "PSE", wb_data$countrycode)
+wb_data$countrycode <- ifelse(wb_data$economy == "Kosovo", "XKX", wb_data$countrycode)
+wb_data$countrycode <- ifelse(wb_data$economy == "Romania", "ROU", wb_data$countrycode)
+wb_data$countrycode <- ifelse(wb_data$economy == "Congo, Dem. Rep.", "COD", wb_data$countrycode)
+wb_data$countrycode <- ifelse(wb_data$economy == "Timor-Leste", "TLS", wb_data$countrycode)
+
+wb_data$economy <- ifelse(wb_data$economy == "Côte d'Ivoire", "Cote d'Ivoire", wb_data$economy)
+wb_data$economy <- ifelse(wb_data$economy == "São Tomé and Príncipe", "Sao Tome and Principe", wb_data$economy)
+wb_data$economy <- ifelse(wb_data$economy == "Puerto Rico (U.S.)", "Sao Tome and Principe", wb_data$economy)
+wb_data$economy <- ifelse(wb_data$economy == "North Macedonia", "Macedonia, FYR", wb_data$economy)
+
+write_csv(
+  x = wb_data,
+  file = "outputs/data/wb_data_clean.csv", 
+)

### 2024-02-18 scripts/03-test_data.R
started writing some tests

@@ -10,6 +10,28 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(readr)
 
-#### Test data ####
+wb_data_clean <- read_csv("outputs/data/wb_data_clean.csv")
+
+#### Run similar tests to simulated data  ####
+
+# Does the year fall between 1971-2020?
+wb_data_clean$Report_Year |> min() == 1971
+wb_data_clean$Report_Year |> max() == 2020
+
+#Does the WBL index stay at or higher than 0?
+wb_data_clean$WBL_Index |> min() == 0
+
+# Do all the index values fall between 0 and 100 (as spec'd in the paper)?
+
+#TODO: fix the c() so that it is a range(0-100)
+
+wb_data_clean$GR1_mobility %in% c(0, 25, 50, 75, 100) 
+wb_data_clean$GR2_workplace %in% c(0, 25, 50, 75, 100)
+wb_data_clean$GR3_pay %in% c(0, 25, 50, 75, 100)
+wb_data_clean$GR4_marriage %in% c(0, 25, 50, 75, 100)
+wb_data_clean$GR5_parenthood %in% c(0, 25, 50, 75, 100)
+wb_data_clean$GR6_entrepreneurship %in% c(0, 25, 50, 75, 100)
+wb_data_clean$GR7_assets %in% c(0, 25, 50, 75, 100)
+wb_data_clean$GR8_pension %in% c(0, 25, 50, 75, 100)

### 2024-02-18 scripts/03-test_data.R
tested data

@@ -1,11 +1,8 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Runs various unit tests on wb_data_clean.csv
+# Author: Luca Carnegie
+# Date: February 18 2024
+# Contact: luca.carnegie@mail.utoronto.ca
 
 
 #### Workspace setup ####
@@ -16,22 +13,23 @@ wb_data_clean <- read_csv("outputs/data/wb_data_clean.csv")
 
 #### Run similar tests to simulated data  ####
 
+# For the specified columns, we test...
+
 # Does the year fall between 1971-2020?
 wb_data_clean$Report_Year |> min() == 1971
 wb_data_clean$Report_Year |> max() == 2020
 
-#Does the WBL index stay at or higher than 0?
-wb_data_clean$WBL_Index |> min() == 0
+# Does the WBL index stay at or higher than 0?
+wb_data_clean$WBL_Index |> min() >= 0
 
-# Do all the index values fall between 0 and 100 (as spec'd in the paper)?
+# Do all the WBL index values fall between 0 and 100 (as spec'd in the paper)?
 
-#TODO: fix the c() so that it is a range(0-100)
+all (wb_data_clean$GR1_mobility >= 0 & wb_data_clean$GR1_mobility <= 100)
+all (wb_data_clean$GR2_workplace >= 0 & wb_data_clean$GR2_workplace <= 100)
+all (wb_data_clean$GR3_pay >= 0 & wb_data_clean$GR3_pay <= 100)
+all (wb_data_clean$GR4_marriage >= 0 & wb_data_clean$GR4_marriage <= 100)
+all (wb_data_clean$GR5_parenthood >= 0 & wb_data_clean$GR5_parenthood <= 100)
+all (wb_data_clean$GR6_entrepreneurship >= 0 & wb_data_clean$GR6_entrepreneurship <= 100)
[…]

### 2024-02-18 scripts/04-play_with_data.R
replicated 1A

@@ -1,0 +1,78 @@
+#### Preamble ####
+# Purpose: Rough code-sketches of replications, etc.
+# Author: Luca Carnegie
+# Date: February 18 2024
+# Contact: luca.carnegie@mail.utoronto.ca
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(readr)
+
+wb_data <- read_csv("outputs/data/wb_data_clean.csv")
+
+
+## Replication of Figure 1A: Avg. WBL Index in 2019
+
+region_avg_index <- wb_data |> group_by(Region) |> filter(`WBL Report Year` == 2019) |>
+                    summarise(avg_index = mean(`WBL_index`, na.rm = TRUE), .groups = "drop")
+
+
+# Plot in ggplot2
+
+  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
+  geom_col() +
+  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
+  theme_minimal() +
+  ylim (NA, 100) +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))
+  
+## Replication of Figure 1B: Indicator variables averages 2019
+
+average_indicator <- wb_data |> 
+                        rename(
+                          Mobility = GR1_mobility, 
+                          Workplace = GR2_workplace, 
+                          Pay = GR3_pay, 
+                          Marriage = GR4_marriage, 
+                          Parenthood = GR5_parenthood, 
+                          Entrepreneurship = GR6_entrep, 
+                          Assets = GR7_assets, 
+                          Pension = GR8_pension
+                        ) |>
+                        filter(`WBL Report Year` == 2019) |>
+                        select(Mobility, 
+                               Workplace, 
+                               Pay, 
+                               Marriage, 
+                               Parenthood, 
+                               Entrepreneurship, 
+                               Assets, 
+                               Pension
+                               ) |>
+                        colMeans()
+
+setNames(cbind(rownames(average_indicator), average_indicator, row.names = NULL), 
+         c("Indicator", "Average WBL Index"))
+
+# Plot in ggplot2
+
+ggplot(data = average_indicator, mapping = aes(x = Region, y = avg_index)) +
+  geom_col() +
[…]

### 2024-02-19 outputs/paper/paper.qmd
finish graphs, edit some of the paper

@@
 ---
-title: "Women and Gender Pay Gap "
-subtitle: "Faulty World Bank Data Conflicts: A Canadian Example"
+title: "The World Bank, Women, Business and the Law"
+subtitle: "The Need for better checking of Data"
 author: 
@@
 thanks: "Code and data are available at: https://github.com/lcarnegie/replicationpaper. "
-date: today
+date: February 20, 2024
 date-format: long
@@
 
-```{r}
+```{r workspace-setup}
 #| include: false
@@
 library(tidyverse)
+
+## WBL Dataset Setup
+
+wb_data <- read_csv(here::here("outputs/data/wb_data_clean.csv"))
 ```
@@
 # Data {#sec-data}
+
+## Source {#subsec1-data}
+## Methodology {#subsec2-data}
 
@@
 
+# References
+
+::: {#refs}
+:::
+
+\newpage
+
 \appendix
@@
 # Appendix {-}
-
-\newpage
-
-# References

### 2024-02-19 outputs/paper/paper.tex
finish graphs, edit some of the paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The World Bank, Women, Business and the Law},
  pdfauthor={Sehar Bajwa; Luca Carnegie},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The World Bank, Women, Business and the Law\thanks{Code and data
are available at: https://github.com/lcarnegie/replicationpaper.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{The Need for better checking of Data}
\author{Sehar Bajwa \and Luca Carnegie}
\date{February 20, 2024}

\begin{document}
\maketitle
\begin{abstract}
First sentence. Second sentence. Third sentence. Fourth sentence.
\end{abstract}

\section{Introduction}\label{introduction}

\section{Data}\label{sec-data}

\subsection{Source}\label{subsec1-data}

\subsection{Methodology}\label{subsec2-data}

\section{Results}\label{sec-results}

\section{Discussion}\label{discussion}

\subsection{First discussion point}\label{sec-point1}

\subsection{Second discussion point}\label{sec-point2}

\subsection{Third discussion point}\label{sec-point3}

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

\newpage

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\end{CSLReferences}

\newpage

\appendix

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}

\end{document}

### 2024-02-19 scripts/04-play_with_data.R

@@ -14,6 +14,8 @@ wb_data <- read_csv("outputs/data/wb_data_clean.csv")
 
 ## Replication of Figure 1A: Avg. WBL Index in 2019
 
+
+# average out the index across regions
 region_avg_index <- wb_data |> group_by(Region) |> filter(`WBL Report Year` == 2019) |>
                     summarise(avg_index = mean(`WBL_index`, na.rm = TRUE), .groups = "drop")
 
@@ -31,6 +33,7 @@ region_avg_index <- wb_data |> group_by(Region) |> filter(`WBL Report Year` == 2
   
 ## Replication of Figure 1B: Indicator variables averages 2019
 
+# isolate indicators 
 average_indicator <- wb_data |> 
                         rename(
                           Mobility = GR1_mobility, 
@@ -51,22 +54,53 @@ average_indicator <- wb_data |>
                                Entrepreneurship, 
                                Assets, 
                                Pension
-                               ) |>
-                        colMeans()
+                               )
 
-setNames(cbind(rownames(average_indicator), average_indicator, row.names = NULL), 
-         c("Indicator", "Average WBL Index"))
+#average them out
+avg_values <- colMeans(average_indicator)
+
+#put them in a table 
+new_table <- data.frame(Column_Names = names(avg_values), Average_Values = avg_values)
+
+
 
 # Plot in ggplot2
 
-ggplot(data = average_indicator, mapping = aes(x = Region, y = avg_index)) +
+  ggplot(data = new_table, mapping = aes(x = Column_Names, y = Average_Values)) +
   geom_col() +
-  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
+  labs(title="Indicator Averages, 2019", x="Indicator", y="WBL Indicator Score") +
   theme_minimal() +
   ylim (NA, 100) +
   theme(plot.title.position = "plot",
         plot.title = element_text(hjust = 0.5), 
         axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))
+  
+  
+## Replication of Figure 2:  Charting the Progress of Legal Gender Equality over Time
+  
+
+#construct tibble of yearly averages of regions.   
+regional_avgs_byyear <- wb_data |> select(
+                            `WBL Report Year`, 
+                            Region, 
+                            WBL_index
+                        ) |> group_by(Region, `WBL Report Year`) |> 
[…]

### 2024-02-19 outputs/paper/paper.qmd
wrote some of paper.

@@
 
+Despite the many decades' worth of progress in gender equality throughout the world, one of the enduring types of gender discrimination worldwide is in the eyes of the law.
+Though much progress has been made in western, high-income countries, there remains vast swaths of the world where women may not vote, work, or participate in the same activities as men do, due to the presence of gendered laws.
+To trace the progression of economic opportunities of women in the world's 190 countries, the World Bank constructed their "Women, Business, and the Law" (WBL) dataset, which Hyland et.
+al make of primary interest in their paper "Gendered Laws and Women in the Workforce".
+In Sections 2 and 3, we analyze and replicate some of the key graphs and insights from their paper.
+Building on their work in Sections 4 and 5, we  call into question the data collection methods of the World Bank, finding data from Statistics Canada that conflict with the data [or whatever Sehar finds], outlining the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.
+
 # Data {#sec-data}
 
+Hyland et al.'s original paper primarily makes use of the 'Women, Business, and the Law' dataset (WBL) from the World Bank, along with some others to help compare and contextualize the data.
+In our replication, however, we will be focusing on aspects of this dataset only.
+
 ## Source {#subsec1-data}
+
+The Women, Business, and the Law dataset is organized across thirty-five aspects of the law, which are scored across eight indicators of four or five binary questions.
+Each indicator represents a different phase of a woman’s career: Mobility, Workplace, Pay, Marriage, Parenthood, Entrepreneurship, Assets, and Pension.
+Answers to the binary questions were sourced, pro-bono, from respondents with expertise in laws on family, labor and violence against women.
+It not clear where in their countries' legal systems the respondents were when they were polled.
+The indicators are as follows:
+
+	Mobility: Examines constraints on women’s freedom of movement. 
+	Workplace: Analyzes laws affecting a woman’s decision to work
+	Pay: Measures laws and regulations affecting a woman’s pay
+	Marriage: Assesses constraints related to marriage
+	Parenthood: Examines laws that affect women’s work after having children
+	Entrepreneurship: Assesses constraints to women starting and running a business
+	Assets: covers property ownership rights, inheritance rights (both for children and surviving spouses), authority of assets during marriage, and valuation of non-monetary contributions. 
+	Pension: captures the equalization of retirement ages (with full and partial benefits as well as the mandatory retirement age) and whether periods of absence from employment due to childcare are accounted for in pension benefits.
+
+	Indicator-level scores are obtained by calculating the unweighted average of the questions within that indicator and scaling the result to 100.
+Overall scores are then calculated by taking the average of each indicator, with 100 representing the highest possible score.
+Each binary question either contributes 20 or 25 points, depending on the indicator.
+The indicators and index mainly capture the effect labour laws have on the employment outcomes of women in the labour force.
+It does not take into account laws affecting their acquisition of human capital (e.g.
+required education laws) nor laws affecting reproductive rights, for instance, which Hyland et al.
+point to as an issue but concede would be difficult to quantify easily.
+The World Bank's methodology was designed as an easily replicable measure of the legal environment for women as entrepreneurs and employees, which, for a dataset of it's size and breadth, is comprehensive.
+In their paper, they then use the dataset to present 'stylized facts' which either unearth new facts about gendered laws or are quantifications of what they write were previously known qualitatively.
+
+	Figures 1A, 1B, and 2 were replicated from the original paper.
+
+```{r hyland-fig-1A}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Figure 1A from Hyland et al. 
+#| fig-align: center
+
+## Replication of Figure 1A: Avg. WBL Index in 2019
+
+# average out the index across regions
+region_avg_index <- wb_data |> group_by(Region) |> filter(`WBL Report Year` == 2019) |>
+                    summarise(avg_index = mean(`WBL_index`, na.rm = TRUE), .groups = "drop")
+
+# Plot in ggplot2
+
+  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
+  geom_col() +
+  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
+  theme_minimal() +
+  ylim (NA, 100) +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))
+
+```
+
+	Figure 1A shows the average unweighted aggregate WBL index score in 2019, compared across regions.
+It shows notable differences between higher-income countries and the rest of the world, most notably between higher income countries and countries in the Middle East and North Africa - there is a almost 25 index-point difference between them.
+In the context of this dataset, it indicates that women in high-income countries enjoy about 36 percent more legal rights related to employment than those in the Middle East and North Africa.
+
+```{r hyland-fig-1B}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Figure 1B from Hyland et al. 
+#| fig-align: center
+
+## Replication of Figure 1A: Avg. WBL Index in 2019
+
+# average out the index across regions
+region_avg_index <- wb_data |> group_by(Region) |> filter(`WBL Report Year` == 2019) |>
+                    summarise(avg_index = mean(`WBL_index`, na.rm = TRUE), .groups = "drop")
+
+# Plot in ggplot2
+
+  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
+  geom_col() +
+  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
+  theme_minimal() +
+  ylim (NA, 100) +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))
+
+```
+
+	Figure 1B breaks down aggregate index for 2019 into it's constituent indicators and shows the global average score for each indicator.
+As Hyland et al pointed out, the 'Parenthood' Indicator was the lowest overall, meaning that globally, women generally enjoy fewer legal rights in employment after they have had children than before.
+
+```{r hyland-fig-2}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Figure 2 from Hyland et al. 
+#| fig-align: center
+
+## Replication of Figure 2:  Charting the Progress of Legal Gender Equality over Time
+
+#construct tibble of yearly averages of regions.   
+regional_avgs_byyear <- wb_data |> select(
+                            `WBL Report Year`, 
+                            Region, 
+                            WBL_index
+                        ) |> group_by(Region, `WBL Report Year`) |> 
+                             summarise(avg_index = mean(`WBL_index`, na.rm = TRUE), .groups = "drop")
+
+# Sling into ggplot2
+
+regional_avgs_byyear |> 
+  ggplot(aes(x=`WBL Report Year`, y=avg_index, group=Region, color=Region)) +
+  geom_line() +
+  labs(title="WBL Index Progression since 1971, by Region", x="Year", y="WBL Index") +
+  scale_color_manual(values=c("red", "blue", "green", "orange", "purple", "turquoise", "black"), labels=c("East Asia/Pacific","Europe/Central Asia", "High Income Countries (OECD)", "Latin America/Caribbean", "Middle East/North Africa", "South Asia", "Sub-Saharan Africa")) +
+  theme_minimal() +
+  ylim (NA, 100) + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+Figure 2 shows the progression of the aggregate WBL index over time, broken down by region.
+As expected, high-income countries dominate in index score, but interestingly Latin American women enjoyed more rights than even European/Central Asian women between 1995-2000.
+Employment rights for women in Sub-Saharan Africa have also seen the most growth since tracking began in 1971.
+
 ## Methodology {#subsec2-data}
+
+R [cite] and it's tidyverse [cite] and dplyr [cite] packages were primarily used when analysing the WBL dataset.
+Like Hyland, this dataset will be of main focus in our analysis.
+Other datasets that helped their paper, such as Wage Gap data from the International Labour Organization and the OECD (Organization for Economic Cooperation and Development), and World Development Indicators data is not important to our analysis.
 

### 2024-02-19 outputs/paper/paper.tex
wrote some of paper.

@@
 
+Despite the many decades' worth of progress in gender equality
+throughout the world, one of the enduring types of gender discrimination
+worldwide is in the eyes of the law. Though much progress has been made
+in western, high-income countries, there remains vast swaths of the
+world where women may not vote, work, or participate in the same
+activities as men do, due to the presence of gendered laws. To trace the
+progression of economic opportunities of women in the world's 190
+countries, the World Bank constructed their ``Women, Business, and the
+Law'' (WBL) dataset, which Hyland et. al make of primary interest in
+their paper ``Gendered Laws and Women in the Workforce''. In Sections 2
+and 3, we analyze and replicate some of the key graphs and insights from
+their paper. Building on their work in Sections 4 and 5, we call into
+question the data collection methods of the World Bank, finding data
+from Statistics Canada that conflict with the data {[}or whatever Sehar
+finds{]}, outlining the importance that accurate data have in being able
+to fully understand the complex picture of women and the law throughout
+the world.
+
 \section{Data}\label{sec-data}
+
+Hyland et al.'s original paper primarily makes use of the `Women,
+Business, and the Law' dataset (WBL) from the World Bank, along with
+some others to help compare and contextualize the data. In our
+replication, however, we will be focusing on aspects of this dataset
+only.
 
@@
 
-\subsection{Methodology}\label{subsec2-data}
+The Women, Business, and the Law dataset is organized across thirty-five
+aspects of the law, which are scored across eight indicators of four or
+five binary questions. Each indicator represents a different phase of a
+woman's career: Mobility, Workplace, Pay, Marriage, Parenthood,
+Entrepreneurship, Assets, and Pension. Answers to the binary questions
+were sourced, pro-bono, from respondents with expertise in laws on
+family, labor and violence against women. It not clear where in their
+countries' legal systems the respondents were when they were polled. The
+indicators are as follows:
+
+\begin{verbatim}
+Mobility: Examines constraints on women’s freedom of movement. 
+Workplace: Analyzes laws affecting a woman’s decision to work
+Pay: Measures laws and regulations affecting a woman’s pay
+Marriage: Assesses constraints related to marriage
+Parenthood: Examines laws that affect women’s work after having children
+Entrepreneurship: Assesses constraints to women starting and running a business
+Assets: covers property ownership rights, inheritance rights (both for children and surviving spouses), authority of assets during marriage, and valuation of nonmonetary contributions. 
+Pension: captures the equalization of retirement ages (with full and partial benefits as well as the mandatory retirement age) and whether periods of absence from employment due to childcare are accounted for in pension benefits.
+
+Indicator-level scores are obtained by calculating the unweighted average of the questions within that indicator and scaling the result to 100.
+Overall scores are then calculated by taking the average of each indicator, with 100 representing the highest possible score.
+Each binary question either contributes 20 or 25 points, depending on the indicator.
+The indicators and index mainly capture the effect labour laws have on the employment outcomes of women in the labour force.
+It does not take into account laws affecting their acquisition of human capital (e.g.
+required education laws) nor laws affecting reproductive rights, for instance, which Hyland et al.
+point to as an issue but concede would be difficult to quantify easily.
+The World Bank's methodology was designed as an easily replicable measure of the legal environment for women as entrepreneurs and employees, which, for a dataset of it's size and breadth, is comprehensive.
+In their paper, they then use the dataset to present 'stylized facts' which either unearth new facts about gendered laws or are quantifications of what they write were previously known qualitatively.
+
+Figures 1A, 1B, and 2 were replicated from the original paper.
+
+Figure 1A shows the average unweighted aggregate WBL index score in 2019, compared across regions.
+It shows notable differences between higher-income countries and the rest of the world, most notably between higher income countries and countries in the Middle East and North Africa - there is a almost 25 index-point difference between them.
+In the context of this dataset, it indicates that women in high-income countries enjoy about 36 percent more legal rights related to employment than those in the Middle East and North Africa.
+
+Figure 1B breaks down aggregate index for 2019 into it's constituent indicators and shows the global average score for each indicator.
+As Hyland et al pointed out, the 'Parenthood' Indicator was the lowest overall, meaning that globally, women generally enjoy fewer legal rights in employment after they have had children than before.
+\end{verbatim}
+
+Figure 2 shows the progression of the aggregate WBL index over time,
+broken down by region. As expected, high-income countries dominate in
+index score, but interestingly Latin American women enjoyed more rights
+than even European/Central Asian women between 1995-2000. Employment
+rights for women in Sub-Saharan Africa have also seen the most growth
+since tracking began in 1971. \#\# Methodology \{\#subsec2-data\}
 

### 2024-02-20 outputs/paper/paper.qmd

@@
 
-# Results {#sec-results}
+# A focus on Canada
 
-# Discussion
+In the subsequent section of this paper, our attention will be directed towards a comprehensive cross-referencing analysis, juxtaposing Canada's standing in the World Bank's Women, Business, and the Law (WBL) index across all eight indicators with research data sourced from Statistics Canada.
+This approach aims to delve deeper into the nuances of Canada's performance in the realm of gender-related legal frameworks and business regulations, providing a nuanced perspective that goes beyond the surface-level rankings.
 
-## First discussion point {#sec-point1}
+ By synthesizing the insights derived from the WBL indicators with the robust statistical data from Statistics Canada, we seek to gain a more thorough understanding of the effectiveness and practical implications of Canada's policies and regulations pertaining to women in the business sector.
+This integrated examination will contribute to a more comprehensive and nuanced evaluation, shedding light on potential areas of alignment or incongruence between the World Bank's assessment and the empirical data provided by Statistics Canada.
+Through this interdisciplinary approach, we aim to offer a more holistic and informed analysis of the Canadian landscape in terms of gender inclusivity within the legal and business domains.
+
+## Canada in the WBL
+
+OECD Countries fare particularly well in the WBL Rankings, as evidenced in the graphs in the former sections, with the highest average WBL Indexes and Index Progressions.
+As a High Income OECD Country, Canada fares particularly well in the WBL Database, scoring 100 across all 8 indicators over the past decade.
+Cross-referencing the updated database with the one used in the paper reveals the same scores till 2023 as well, revealing that not much has changed post the COVID 19 Pandemic.
+
+Every indicator within the study is supported by a 'legal basis' as outlined in the original WBL Database, providing a foundation to substantiate the assigned marks for each question.
+It is essential to note that in the datasheet utilized by Hyland et al, the legal columns were omitted.
+However, for the purpose of our subsequent analysis, we deem it imperative to reference the official database's legal columns.
+This is crucial as they not only serve to validate our findings but also contribute valuable insights that may be overlooked when solely relying on the abbreviated dataset used by Hyland et al.
+
+# Analysing WBL Indicators from a Canadian viewpoint
+
+## Pay
+
+The first indicator to be scrutinized is Pay, particularly the question that focuses on equal remuneration regardless of gender.
+Notably, the legal backing for an indicator representative of Canada is one that references only one province, Ontario.
+Nevertheless, this section will utilize the LFS Survey from Statistics Canada to evaluate if Canada’s perfect rating doesn’t hold up solely on paper.
+
+```{r, results='asis', echo=FALSE}
+library(kableExtra)
+
+# Create a data frame for the table
+data <- data.frame(
+  Indicator = "Pay",
+  Score = 100,
+  Question = "“Does the law mandate equal remuneration for work of equal value?”",
+  Legal_Basis = "Ontario Pay Equity Act, Sec. 6",
+  Achievement_of_pay_equity = "6 (1) For the purposes of this Act, pay equity is achieved under the job-to-job method of comparison when the job rate for the female job class that is the subject of the comparison is at least equal to the job rate for a male job class in the same establishment where the work performed in the two job classes is of equal or comparable value.
+R.S.O.
+1990, c.
+P.7, s.
+6 (1); 1993, c.
+4, s.
+4 (1).",
+  Relevant_article = "The gender wage gap in Canada: 1998 to 2018 C Labour Statistics: Research Papers by Rachelle Pelletier, Martha Patterson and Melissa Moyser",
+  Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
+  Link_to_Article = "[Read more](https://www150.statcan.gc.ca/n1/en/pub/75-004-m/75-004-m2019004-eng.pdf?st=szPFyv6s)",
+  Database = "Annualized data from the Labour Force Survey (LFS), a monthly household survey conducted by Statistics Canada.",
+  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
+)
+
+# Transpose the data frame
+data <- t(data)
+
+# Create a table using kableExtra
+kable(data, format = "markdown") %>%
+  kable_styling()
+```
 

### 2024-02-20 outputs/paper/paper.qmd
Adding workplace and pay

@@
 # Create a data frame for the table
-data <- data.frame(
+pay <- data.frame(
   Indicator = "Pay",
@@
 # Transpose the data frame
-data <- t(data)
+pay <- t(pay)
+
+# Replace underscores with spaces in column names
+colnames(pay) <- gsub("_", " ", colnames(pay))
 
 # Create a table using kableExtra
-kable(data, format = "markdown") %>%
-  kable_styling()
+  kable(pay, format = "latex", caption = "Summary of findings -Pay Indicator") %>%
+  column_spec(2, width = "27em")
+
 ```
 
-## Second discussion point {#sec-point2}
+The LFS survey reveals that in 2018, female employees aged 25 to 34 earned $4.13 less per hour compared to their male counterparts, equaling 87 cents on the dollar.
+In the decade between 1988 and 2018, the gap has reduced by merely $1.04.
+Given that women in Canada are now better educated on average than men, acquiring bachelor’s degrees at a faster pace, and have diversified their fields of study, particularly in STEM, the persistence of this gap warrants a closer look.
 
-## Third discussion point {#sec-point3}
+The authors used a Blinder Oaxaca Decomposition to quantify how much of the wage difference can be explained by control factors.
+Some of these factors included level of education, job tenure, union coverage, and demographics.
+The results show that even after controlling for the above factors, roughly two-thirds of the wage gap is unexplained, and the difference is statistically significant at a 95% confidence interval.
+
+The table below reproduces the gender wage gap- Oaxaca decomposition data from the Labour Statistics report.
+As is visible, only an additional 7% of this gap has become explicable in the years that have followed since 1988, largely due to unionization reducing men's wages, and women entering new industries.
+Despite extensive research, the gender wage gap remains an unsolved mystery in Canada.
+
+```{r, results='asis', echo=FALSE, fig.cap="Explanation of the Gender Wage Gap, 1998 to 2018"}
+
+library(kableExtra)
+
+# Create a data frame for the table
+wagegap <- data.frame(
+  Category = c("Total explained portion (sum of effects of variables below)", "Human capital", "Education", 
+               "Job tenure", "Job attributes", "Part-time", "Public sector", "Union", "Firm size", 
+               "Occupation and industry", "Occupation", "Industry", "Demographics", "Age", 
+               "Province", "Children", "Marital status", "Total unexplained portion"),
+  'In 1998' = c(28.0, 1.8, -0.5, 2.3, 10.1, 8.9, -0.6, 0.7, 1.1, 18.3, 1.8, 16.5, -2.2, -0.4, -0.1, -0.3, -1.4, 72.0),
+  'In 2018' = c(36.6, -6.1, -4.8, -1.3, -0.7, 9.2, -5.3, -3.4, -1.1, 44.8, 5.1, 39.7, -1.6, -0.3, 1.0, -0.8, -1.5, 63.4)
+)
+
+# Replace special characters in column names
+#colnames(wagegap) <- gsub("`", "", colnames(wagegap))
+
+# Create a table using kableExtra
+kable(wagegap, format = "latex", booktabs = TRUE, escape = FALSE) %>%
+ kable_styling() %>%
+#  add_header_above(c("Category" = 1, "1998" = 1, "2018" = 1)) %>%
+  row_spec(18, bold = TRUE)
+
+```
+
+## Workplace {#sec-point2}
+
+The next indicator under scrutiny is the Workplace, with a specific focus on legislation addressing sexual harassment.
+While legal support is evident in various codes within this category, it is essential to emphasize that the mere existence of legal provisions does not guarantee their effective implementation.
+The following analysis then, acknowledges the potential disparity between legal frameworks and their real-world impact on preventing and addressing sexual harassment in the workplace.
+
+```{r, results='asis', echo=FALSE}
+library(kableExtra)
+
+# Create a data frame for the table
+workplace <- data.frame(
+  Indicator = "Workplace",
+  Score = 100,
+  Question = "“Is there legislation on sexual harassment in employment??”",
+  Legal_Basis = "Canadian Human Rights Act, Sec. 14(1) and (2); Ontario Occupational Health and Safety Act, Sec. 1 and 32.0.1; Ontario Human Rights Code, Sec. 7(2) and 7(3)",
+  Code = "14 (1) It is a discriminatory practice, in matters related to employment,
+to harass an individual on a prohibited ground of discrimination.Without limiting the generality of subsection 1, sexual harassment shall, for the purposes of that subsection, be deemed to be harassment on a prohibited ground of discrimination.",
+  Relevant_article = "Gender Results Framework: A new data table on workplace harassment",
+  Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
+  Link_to_Article = "[Read more]https://www150.statcan.gc.ca/n1/daily-quotidien/240212/dq240212a-eng.htm",
+  Database = "Data from 2020 Survey on Sexual Misconduct at Work (SSMW) and Gender Results Framework",
+  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
+)
+
+# Transpose the data frame
+workplace <- t(workplace)
+
+# Replace underscores with spaces in column names
+colnames(workplace) <- gsub("_", " ", colnames(workplace))
+
+# Create a table using kableExtra
+  kable(workplace, format = "latex", caption = "Summary of findings -Pay Indicator") %>%
+  column_spec(2, width = "27em")
+
+```
+
+Utilizing data from Statistics Canada's Centre for Gender, Diversity, and Inclusion Statistics, in collaboration with the Canadian Centre for Justice and Community Safety Statistics, we analyze a recently released data table on workplace harassment and sexual assault experiences.
+The table reports on the Gender Results Framework (GRF), initiated by the Government of Canada in 2018 to monitor gender equality progress.
+Additionally, the data incorporates insights gathered from the 2020 Survey on Sexual Misconduct at Work (SSMW).
+
+The reproduced table from the 2020 Survey on Sexual Misconduct at Work outlines the proportions of employees reporting inappropriate sexualized behavior, discriminatory behavior, and sexual assault in their work environments.
+Notably, the data highlights the heightened vulnerability of individuals in the 25 to 34 age group, where 57% of females report experiencing sexualized behaviors.
+Within this age group, women are twice as likely as men to face discrimination and three times as likely to experience sexual assault.
+```{r}
+# Install and load necessary packages if not already installed
+ # install.packages("tidyverse")
+
+# Install and load necessary packages if not already installed
+if (!requireNamespace("tidyverse", quietly = TRUE)) {
+  install.packages("tidyverse")
+}
+
+# Load the necessary libraries
+library(tidyverse)
+
+# Your provided data
+data <- read.table(text = '"Group" "M - inappropriate sexualized behaviours" "M - Discrimination" "M - SA" "F - inappropriate sexualized behaviours" "F - Discrimination" "F - SA"
+"Total, all people" 29 9 3 44 20 13
+"Age 15 to 24" 27 12 0 48 25 7
+"Age 25 to 34" 37 13 5 57 27 16
+"Age 35 to 44" 34 11 3 41 20 12
+"Age 45 to 54" 28 8 3 42 18 12
+"Age 55 and older" 24 6 2 39 16 15', header = TRUE)
+
+# Transform the data into long format for ggplot
+data_long <- data %>%
+  pivot_longer(cols = -Group, names_to = "Category", values_to = "Percentage")
+
+# Create a sideways stacked bar plot
+ggplot(data_long, aes(x = Percentage, y = Group, fill = Category)) +
+  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
+  scale_fill_manual(values = c("#FF9999", "#FF6666", "#FF3333", "#66B2FF", "#3366FF", "#000099"), name = "Category") +
+  labs(title = "Sideways Stacked Bar Plot",
+       x = "Percentage",
+       y = "Group") +
+  theme_minimal() +
+  theme(
+    axis.text.y = element_text(angle = 0, hjust = 1),
+    legend.position = "bottom",
+    legend.direction = "horizontal",
+    legend.box = "horizontal",
+    legend.margin = margin(t = 0, unit = "pt")
+)
+
+```
+
+Further analysis of Canadian data reveals that nearly half of women, at 47%, report experiencing some form of harassment or sexual assault in the workplace.
+The Gender Results Framework (GRF), a key tool for monitoring gender equality in Canada, indicates that women are twice as likely as men to encounter unwanted sexual behaviors in a work environment.
+Beyond the workplace, Statistics Canada notes that women are three times more likely to experience unwanted sexual behaviors in public, four times more likely to have been sexually assaulted since age 15, and twice as likely to suffer negative emotional consequences from such experiences compared to men.
+These findings underscore the pervasive nature of gender-based harassment and assault, both within and outside professional settings in Canada, which are far from a pristine perfect ‘100'.
 

### 2024-02-20 outputs/paper/paper.qmd
changes to graphs

@@
 
-```{r, results='asis', echo=FALSE}
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Figure for pay
+#| fig-align: center
   library(kableExtra)
@@
 ```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Figure 
+#| fig-align: center
 # Install and load necessary packages if not already installed
@@
 )
-
 ```

### 2024-02-21 outputs/paper/paper.qmd
updated references

--- 2024-02-20 outputs/paper/paper.qmd
@@
 In Sections 2 and 3, we analyze and replicate some of the key graphs and insights from their paper.
-Building on their work in Sections 4 and 5, we  call into question the data collection methods of the World Bank, finding data from Statistics Canada that conflict with the data [or whatever Sehar finds], outlining the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.
+Building on their work in Sections 4 and 5, we call into question the data collection methods of the World Bank, finding data from Statistics Canada that conflict with the data \[or whatever Sehar finds\], outlining the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.
 
-# Data {#sec-data}
+# Data
+
+## Source {#sec-data}
 
@@
 
-## Source {#subsec1-data}
+## Methodology {#subsec2-data}
+
+`R` [@citeR] and it's `tidyverse` [@tidyverse] and `dplyr` [@dplyr] packages were primarily used when analysing the WBL dataset.
+Like Hyland, this WBL dataset will be of primary focus in our reproduction.
+Other datasets that informed the original paper and were cross-references, such as Wage Gap data from the International Labour Organization and the OECD (Organization for Economic Cooperation and Development), and World Development Indicators data are not relevant to the areas that will be developed upon, and so irrelevant to the subsequent analysis.
+For the secondary part of this analysis, Surveys conducted by Statistics Canada will be utilised.
+
+## Features {#subsec1-data}
 
@@
 
-	Mobility: Examines constraints on women’s freedom of movement. 
-	Workplace: Analyzes laws affecting a woman’s decision to work
-	Pay: Measures laws and regulations affecting a woman’s pay
-	Marriage: Assesses constraints related to marriage
-	Parenthood: Examines laws that affect women’s work after having children
-	Entrepreneurship: Assesses constraints to women starting and running a business
-	Assets: covers property ownership rights, inheritance rights (both for children and surviving spouses), authority of assets during marriage, and valuation of non-monetary contributions. 
-	Pension: captures the equalization of retirement ages (with full and partial benefits as well as the mandatory retirement age) and whether periods of absence from employment due to childcare are accounted for in pension benefits.
+• Mobility: Examines constraints on women’s freedom of movement.\
+• Workplace: Analyzes laws affecting a woman’s decision to work.\
+• Pay: Measures laws and regulations affecting a woman’s pay.\
+• Marriage: Assesses constraints related to marriage\
+• Parenthood: Examines laws that affect women’s work after having children.\
+• Entrepreneurship: Assesses constraints to women starting and running a business.\
+• Assets: covers property ownership rights, inheritance rights (both for children and surviving spouses), authority of assets during marriage, and valuation of non-monetary contributions.\
+• Pension: captures the equalization of retirement ages (with full and partial benefits as well as the mandatory retirement age) and whether periods of absence from employment due to childcare are accounted for in pension benefits.
 
@@
 In their paper, they then use the dataset to present 'stylized facts' which either unearth new facts about gendered laws or are quantifications of what they write were previously known qualitatively.
+Figures 1A, 1B, and 2 int the following sections were replicated from the original paper.
 
-	Figures 1A, 1B, and 2 were replicated from the original paper.
+## Key Results
 
@@
 
-## Methodology {#subsec2-data}
-
-R [cite] and it's tidyverse [cite] and dplyr [cite] packages were primarily used when analysing the WBL dataset.
-Like Hyland, this dataset will be of main focus in our analysis.
-Other datasets that helped their paper, such as Wage Gap data from the International Labour Organization and the OECD (Organization for Economic Cooperation and Development), and World Development Indicators data is not important to our analysis.
-
 # A focus on Canada
@@
 
-# Analysing WBL Indicators from a Canadian viewpoint
-
-## Pay
+### Pay
 
@@
 Nevertheless, this section will utilize the LFS Survey from Statistics Canada to evaluate if Canada’s perfect rating doesn’t hold up solely on paper.
-
-```{r, results='asis', echo=FALSE}
+```{r canada-tbl-1}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
 library(kableExtra)
@@
   Legal_Basis = "Ontario Pay Equity Act, Sec. 6",
-  Achievement_of_pay_equity = "6 (1) For the purposes of this Act, pay equity is achieved under the job-to-job method of comparison when the job rate for the female job class that is the subject of the comparison is at least equal to the job rate for a male job class in the same establishment where the work performed in the two job classes is of equal or comparable value.
-R.S.O.
-1990, c.
-P.7, s.
-6 (1); 1993, c.
-4, s.
-4 (1).",
-  Relevant_article = "The gender wage gap in Canada: 1998 to 2018 C Labour Statistics: Research Papers by Rachelle Pelletier, Martha Patterson and Melissa Moyser",
+  Citation = "For the purposes of this Act, pay equity is achieved under the job-to-job method of comparison when the job rate for the female job class that is the subject of the comparison is at least equal to the job rate for a male job class in the same establishment where the work performed in the two job classes is of equal or comparable value.",
+  Article = "The gender wage gap in Canada: 1998 to 2018 Labour Statistics: Research Papers by Rachelle Pelletier, Martha Patterson and Melissa Moyser",
   Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
-  Link_to_Article = "[Read more](https://www150.statcan.gc.ca/n1/en/pub/75-004-m/75-004-m2019004-eng.pdf?st=szPFyv6s)",
+  Link = "[Read more](https://www150.statcan.gc.ca/n1/en/pub/75-004-m/75-004-m2019004-eng.pdf?st=szPFyv6s)",
   Database = "Annualized data from the Labour Force Survey (LFS), a monthly household survey conducted by Statistics Canada.",
-  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
+  Additional = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
 )
@@
   kable(pay, format = "latex", caption = "Summary of findings -Pay Indicator") %>%
-  column_spec(2, width = "27em")
-
+  column_spec(2, width = "29em")
 ```
 
-The LFS survey reveals that in 2018, female employees aged 25 to 34 earned $4.13 less per hour compared to their male counterparts, equaling 87 cents on the dollar.
-In the decade between 1988 and 2018, the gap has reduced by merely $1.04.
+The LFS survey reveals that in 2018, female employees aged 25 to 34 earned 4.13 dollars less per hour compared to their male counterparts, equaling 87 cents on the dollar.
+In the decade between 1988 and 2018, the gap has reduced by merely \$1.04.
 Given that women in Canada are now better educated on average than men, acquiring bachelor’s degrees at a faster pace, and have diversified their fields of study, particularly in STEM, the persistence of this gap warrants a closer look.
@@
 
-```{r, results='asis', echo=FALSE, fig.cap="Explanation of the Gender Wage Gap, 1998 to 2018"}
-
+```{r canada-tbl-2}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
 library(kableExtra)
@@
                "Province", "Children", "Marital status", "Total unexplained portion"),
-  'In 1998' = c(28.0, 1.8, -0.5, 2.3, 10.1, 8.9, -0.6, 0.7, 1.1, 18.3, 1.8, 16.5, -2.2, -0.4, -0.1, -0.3, -1.4, 72.0),
-  'In 2018' = c(36.6, -6.1, -4.8, -1.3, -0.7, 9.2, -5.3, -3.4, -1.1, 44.8, 5.1, 39.7, -1.6, -0.3, 1.0, -0.8, -1.5, 63.4)
+  '1998' = c(28.0, 1.8, -0.5, 2.3, 10.1, 8.9, -0.6, 0.7, 1.1, 18.3, 1.8, 16.5, -2.2, -0.4, -0.1, -0.3, -1.4, 72.0),
+  '2018' = c(36.6, -6.1, -4.8, -1.3, -0.7, 9.2, -5.3, -3.4, -1.1, 44.8, 5.1, 39.7, -1.6, -0.3, 1.0, -0.8, -1.5, 63.4)
 )
+
+colnames(wagegap) <- c("Category", "1998", "2018")
 
@@
 # Create a table using kableExtra
-kable(wagegap, format = "latex", booktabs = TRUE, escape = FALSE) %>%
+kable(wagegap, format = "latex", booktabs = TRUE, escape = FALSE, caption = "Explanation of the Gender Wage Gap, 1998 to 2018") %>%
  kable_styling() %>%
 #  add_header_above(c("Category" = 1, "1998" = 1, "2018" = 1)) %>%
+
   row_spec(18, bold = TRUE)
@@
 
-## Workplace {#sec-point2}
+### Workplace {#sec-point2}
 
@@
 
-```{r, results='asis', echo=FALSE}
+```{r canada-tbl-3}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
 library(kableExtra)
@@
   Score = 100,
-  Question = "“Is there legislation on sexual harassment in employment??”",
-  Legal_Basis = "Canadian Human Rights Act, Sec. 14(1) and (2); Ontario Occupational Health and Safety Act, Sec. 1 and 32.0.1; Ontario Human Rights Code, Sec. 7(2) and 7(3)",
-  Code = "14 (1) It is a discriminatory practice, in matters related to employment,
-to harass an individual on a prohibited ground of discrimination.Without limiting the generality of subsection 1, sexual harassment shall, for the purposes of that subsection, be deemed to be harassment on a prohibited ground of discrimination.",
-  Relevant_article = "Gender Results Framework: A new data table on workplace harassment",
+  Question = "“Is there legislation on sexual harassment in employment?”",
+  Legal_Basis = "Canadian Human Rights Act, Sec. 14(1) and (2)",
+  Citation = "It is a discriminatory practice,in matters related to employment,to harass an individual on a prohibited ground of discrimination.",
+  Article = "Gender Results Framework: A new data table on workplace harassment",
   Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
-  Link_to_Article = "[Read more]https://www150.statcan.gc.ca/n1/daily-quotidien/240212/dq240212a-eng.htm",
-  Database = "Data from 2020 Survey on Sexual Misconduct at Work (SSMW) and Gender Results Framework",
-  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
+  Link = "[Read more](https://www150.statcan.gc.ca/n1/daily-quotidien/240212/dq240212a-eng.htm)",
+  Database = "Data from 2020 Survey on Sexual Misconduct at Work (SSMW) and Gender Results Framework."
 )
@@
 colnames(workplace) <- gsub("_", " ", colnames(workplace))
+# Create a table using kableExtra
+  kable(workplace, format = "latex", caption = "Summary of findings -Workplace Indicator") %>%
 
-# Create a table using kableExtra
-  kable(workplace, format = "latex", caption = "Summary of findings -Pay Indicator") %>%
-  column_spec(2, width = "27em")
-
+  column_spec(2, width = "29em")
 ```
@@
 Within this age group, women are twice as likely as men to face discrimination and three times as likely to experience sexual assault.
-```{r}
+
+```{r canada-fig-1}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
 # Install and load necessary packages if not already installed
  # install.packages("tidyverse")
-
-# Install and load necessary packages if not already installed
-if (!requireNamespace("tidyverse", quietly = TRUE)) {
-  install.packages("tidyverse")
-}
 
@@
 
-# Your provided data
+#  provided data
 data <- read.table(text = '"Group" "M - inappropriate sexualized behaviours" "M - Discrimination" "M - SA" "F - inappropriate sexualized behaviours" "F - Discrimination" "F - SA"
@@
 ggplot(data_long, aes(x = Percentage, y = Group, fill = Category)) +
-  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
+  geom_bar(stat = "identity", position = "dodge", width = 1) +
   scale_fill_manual(values = c("#FF9999", "#FF6666", "#FF3333", "#66B2FF", "#3366FF", "#000099"), name = "Category") +
-  labs(title = "Sideways Stacked Bar Plot",
-       x = "Percentage",
-       y = "Group") +
+  labs(title = "Inappropriate workplace behaviour ",
+       x = "Percentage affected ",
+       y = "Age Group") +
+
   theme_minimal() +
   theme(
-    axis.text.y = element_text(angle = 0, hjust = 1),
+ #   axis.text.y = element_text(angle = 0, hjust = 2),
     legend.position = "bottom",
-    legend.direction = "horizontal",
+    legend.direction = "vertical",
     legend.box = "horizontal",
-    legend.margin = margin(t = 0, unit = "pt")
 )
@@
 
-## Weaknesses and next steps
+### Parenthood
+
+```{r canada-tbl-4}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+library(kableExtra)
+
+# Create a data frame for the table
+maternal <- data.frame(
+  Indicator = "Parenthood",
+  Score = 100,
+  Question = "“Is there paid parental leave?”",
+  Legal_Basis = "Employment Insurance Act, Art. 12(4); Employment Insurance Act, Secs. 12(4.01) and 23",
+  Citation = "The maximum number of weeks for which benefits may be paid for a single pregnancy is 15",
+  Article = "2021 MATERNITY LEAVE EXPERIENCE REPORT, Moms At Work",
+  Stylized_fact = "One in every 3 mothers returning to the workplace experienced discrimination",
+  Link = "[Read more](https://thisismomsatwork.com/wp-content/uploads/edd/2021/05/2021-Maternity-Leave-Experience-Report-Moms-at-Work-1.pdf)",
+  Database = "Data from 2021 survey of 1090 mothers who had taken maternity leave in the last 10 years."
+)
+
+# Transpose the data frame
+maternal <- t(maternal)
+
+# Replace underscores with spaces in column names
+colnames(maternal) <- gsub("_", " ", colnames(maternal))
+# Create a table using kableExtra
+  kable(maternal, format = "latex", caption = "Summary of findings -Parenthood Indicator") %>%
+
+  column_spec(2, width = "29em")
+```
+
+This last indicator examines the challenges faced by new mothers returning to the workplace.
+Canada has legal provisions to ensure that maternal leave comes with benefits, and national statistics indicate that a substantial majority, 85.4% of new mothers with insurable employment, received maternity or parental benefits in 2018 according to the Employment Insurance Coverage Survey.
+However, it is crucial to consider the practical implementation of these laws, as their existence doesn't guarantee enforcement.
+A survey conducted among 1090 mothers who took maternity leave between 2011 and 2021 reveals that one in three women encountered discrimination upon their return to the workplace.
+This discrimination manifested in various forms, including the following:\
+• Poor treatment and derogatory comments\
+• Denied opportunities to develop professionally (projects, training, responsibility)\
+• Laid off, restructured or demoted\
+• Denied promotion, new role or advancement of earnings\
+• Inadequate or denial of accommodations.
+
+Furthermore, a striking 95% of the surveyed women did not receive any formal support during their maternity leave transition.
+Notably, 40% seriously contemplated resigning during the return-to-work process.
+
+# Discussion
+
+## Concluding remarks
+
+In conclusion, the Canadian lens applied to the WBL Database provide a
+
+## Rigged data and the World Bank
+
+## Future research
 
@@
 ::: {#refs}
-:::
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}

### 2024-02-21 outputs/paper/paper.qmd
Finished references, wrote abstract, fixed intro

@@
 title: "The World Bank, Women, Business and the Law"
-subtitle: "The Need for better checking of Data"
+subtitle: "How Canada's working women fare in the business world"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Gender-based legal frameworks and workplace conditions are crucial considerations in the pursuit of global gender equality.
+This paper replicates and extends the data analysis conducted by Hyland et al.
+on the World Bank's 'Women, Business, and the Law' dataset.
+By cross-referencing the World Bank's indicators with Statistics Canada's data, we delve into the nuanced realities of gender-related legal frameworks and business regulations.
+The revelations underscore the need for a comprehensive and more practical approach in evaluating gender inclusivity within legal and business domains.
+"
 format: pdf
@@
 bibliography: references.bib
+toc: true
 ---
@@
 
-Despite the many decades' worth of progress in gender equality throughout the world, one of the enduring types of gender discrimination worldwide is in the eyes of the law.
+Despite the many decades' worth of progress in gender equality throughout the world, one of the enduring types of gender discrimination worldwide is in the eyes of the Law.
 Though much progress has been made in western, high-income countries, there remains vast swaths of the world where women may not vote, work, or participate in the same activities as men do, due to the presence of gendered laws.
-To trace the progression of economic opportunities of women in the world's 190 countries, the World Bank constructed their "Women, Business, and the Law" (WBL) dataset, which Hyland et.
-al make of primary interest in their paper "Gendered Laws and Women in the Workforce".
-In Sections 2 and 3, we analyze and replicate some of the key graphs and insights from their paper.
-Building on their work in Sections 4 and 5, we call into question the data collection methods of the World Bank, finding data from Statistics Canada that conflict with the data \[or whatever Sehar finds\], outlining the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.
+To trace the progression of economic opportunities of women in the world's 190 countries, the World Bank constructed their "Women, Business, and the Law" (WBL) dataset [@world_bank], which Hyland et.
+al make of primary interest in their paper "Gendered Laws and Women in the Workforce"[@hylandetal].
+In Sections 2, we analyze and replicate some of the key graphs and insights from their paper.
+Building on their work in Sections 4, we call into question the surface level ratings attributed by the World Bank by cross-referencing with the on-ground realities of gender-based discrimination using data from Statistics Canada.
+The concluding section outlines the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.
 
@@
 In their paper, they then use the dataset to present 'stylized facts' which either unearth new facts about gendered laws or are quantifications of what they write were previously known qualitatively.
-Figures 1A, 1B, and 2 int the following sections were replicated from the original paper.
+@fig-firsthyland, @fig-secondhyland, and @fig-thirdhyland in the following sections were replicated from the original paper.
 
@@
 
-```{r hyland-fig-1A}
+```{r}
+#| label: fig-firsthyland
 #| message: false
@@
 
-Figure 1A shows the average unweighted aggregate WBL index score in 2019, compared across regions.
+@fig-firsthyland shows the average unweighted aggregate WBL index score in 2019, compared across regions.
 It shows notable differences between higher-income countries and the rest of the world, most notably between higher income countries and countries in the Middle East and North Africa - there is a almost 25 index-point difference between them.
@@
 
-```{r hyland-fig-1B}
+```{r}
+#| label: fig-secondhyland
 #| message: false
@@
 
-Figure 1B breaks down aggregate index for 2019 into it's constituent indicators and shows the global average score for each indicator.
+@fig-secondhyland breaks down aggregate index for 2019 into it's constituent indicators and shows the global average score for each indicator.
 As Hyland et al pointed out, the 'Parenthood' Indicator was the lowest overall, meaning that globally, women generally enjoy fewer legal rights in employment after they have had children than before.
 
-```{r hyland-fig-2}
+```{r}
+#| label: fig-thirdhyland
 #| message: false
@@
 
-Figure 2 shows the progression of the aggregate WBL index over time, broken down by region.
+@fig-thirdhyland shows the progression of the aggregate WBL index over time, broken down by region.
 As expected, high-income countries dominate in index score, but interestingly Latin American women enjoyed more rights than even European/Central Asian women between 1995-2000.
@@
 OECD Countries fare particularly well in the WBL Rankings, as evidenced in the graphs in the former sections, with the highest average WBL Indexes and Index Progressions.
-As a High Income OECD Country, Canada fares particularly well in the WBL Database, scoring 100 across all 8 indicators over the past decade.
+As a High Income OECD Country, Canada fares particularly well in the WBL Database [@world_bank], scoring 100 across all 8 indicators over the past decade.
 Cross-referencing the updated database with the one used in the paper reveals the same scores till 2023 as well, revealing that not much has changed post the COVID 19 Pandemic.
@@
 Notably, the legal backing for an indicator representative of Canada is one that references only one province, Ontario.
-Nevertheless, this section will utilize the LFS Survey from Statistics Canada to evaluate if Canada’s perfect rating doesn’t hold up solely on paper.
+Nevertheless, this section will utilize the LFS Survey from Statistics Canada[@wagegap_canada] to evaluate if Canada’s perfect rating doesn’t hold up solely on paper.
 ```{r canada-tbl-1}
@@
 The LFS survey reveals that in 2018, female employees aged 25 to 34 earned 4.13 dollars less per hour compared to their male counterparts, equaling 87 cents on the dollar.
-In the decade between 1988 and 2018, the gap has reduced by merely \$1.04.
+In the decade between 1988 and 2018, the gap has reduced by merely $1.04.
 Given that women in Canada are now better educated on average than men, acquiring bachelor’s degrees at a faster pace, and have diversified their fields of study, particularly in STEM, the persistence of this gap warrants a closer look.
@@
 Some of these factors included level of education, job tenure, union coverage, and demographics.
-The results show that even after controlling for the above factors, roughly two-thirds of the wage gap is unexplained, and the difference is statistically significant at a 95% confidence interval.
+The results in @tbl-explanation show that even after controlling for the above factors, roughly two-thirds of the wage gap is unexplained, and the difference is statistically significant at a 95% confidence interval.
 
@@
 
-```{r canada-tbl-2}
+```{r}
 #| message: false
+#| label: tbl-explanation
 #| echo: false
@@
 
-Utilizing data from Statistics Canada's Centre for Gender, Diversity, and Inclusion Statistics, in collaboration with the Canadian Centre for Justice and Community Safety Statistics, we analyze a recently released data table on workplace harassment and sexual assault experiences.
+Utilizing data from Statistics Canada's Centre for Gender, Diversity, and Inclusion Statistics, in collaboration with the Canadian Centre for Justice and Community Safety Statistics [@harassment_canada], we analyze a recently released data table on workplace harassment and sexual assault experiences.
 The table reports on the Gender Results Framework (GRF), initiated by the Government of Canada in 2018 to monitor gender equality progress.
-Additionally, the data incorporates insights gathered from the 2020 Survey on Sexual Misconduct at Work (SSMW).
+Additionally, the data incorporates insights gathered from the 2020 Survey on Sexual Misconduct at Work (SSMW)[@SSMW].
 
@@
 This last indicator examines the challenges faced by new mothers returning to the workplace.
-Canada has legal provisions to ensure that maternal leave comes with benefits, and national statistics indicate that a substantial majority, 85.4% of new mothers with insurable employment, received maternity or parental benefits in 2018 according to the Employment Insurance Coverage Survey.
+Canada has legal provisions to ensure that maternal leave comes with benefits, and national statistics indicate that a substantial majority, 85.4% of new mothers with insurable employment, received maternity or parental benefits in 2018 according to the Employment Insurance Coverage Survey [@insurance].
 However, it is crucial to consider the practical implementation of these laws, as their existence doesn't guarantee enforcement.
-A survey conducted among 1090 mothers who took maternity leave between 2011 and 2021 reveals that one in three women encountered discrimination upon their return to the workplace.
+A survey conducted among 1090 mothers who took maternity leave between 2011 and 2021 reveals that one in three women encountered discrimination upon their return to the workplace [@moms].
 This discrimination manifested in various forms, including the following:\
@@
 
-In conclusion, the Canadian lens applied to the WBL Database provide a
+In conclusion, the Canadian lens applied to the WBL Database provides a nuanced perspective on the state of gender-related legal frameworks and business regulations.
+While the World Bank's Women, Business, and the Law (WBL) index assigns Canada a perfect score across all eight indicators, our analysis using Statistics Canada data reveals a more complex reality.
 
@@
 
+An inquiry into the manipulation of an annual World Bank report revealed in 2021 that Kristalina Georgieva, the former Chief Executive of the bank and current head of the International Monetary Fund, instructed staff to modify data to appease China.
+The investigation, conducted by the law firm WilmerHale at the request of the bank’s ethics committee,  raised concerns about Georgieva's judgment during her tenure at the World Bank.
+
+In light of these revelations, Luiz Vieira, coordinator of the London-based Bretton Woods Project, a nonprofit watchdog group said:
+"It highlights the degree to which the World Bank and the IMF can really be trusted to provide advice based on solid research.
+It raises questions about whose interests are being served, how robust is their analysis, and how subject to geopolitical and shareholder pressure are they?"
+
+This brings into question the validity of data, even that which is considered unimpeachable and global. One incident of rigged data brought to light underscores the possibility of others.
+
 ## Future research
+
+The closer inspection of Canada in the WBL Database reveals several gaps that could influence future research.
+Principal amongst them is establishing the basis on which legal frameworks are selected to score the countries: as is evidenced above, sometimes codes applying to Ontario are applied, and wherever convenient, National Laws are applied.
+
+Secondly, as is emphasised numerous times in this reproduction, binary questions about whether a law exist or not are just the beginning.
+They do not imply on ground realities, or legal enforcement.
+Then, if the question being asked is "Are women paid the same amount as men?", the answer is a resounding no.
+A useful avenue of research would be to collate from national surveys and frameworks the what the gender-based discrimination situation is in every country.
 
@@
 # References
-
-::: {#refs}
-
-\newpage
-
-\appendix
-
-# Appendix {.unnumbered}

### 2024-02-21 outputs/paper/paper.qmd
fixed graph

@@
 al make of primary interest in their paper "Gendered Laws and Women in the Workforce"[@hylandetal].
-In Sections 2, we analyze and replicate some of the key graphs and insights from their paper.
-Building on their work in Sections 4, we call into question the surface level ratings attributed by the World Bank by cross-referencing with the on-ground realities of gender-based discrimination using data from Statistics Canada.
+In Section 2, we analyze and replicate some of the key graphs and insights from their paper.
+Building on their work in Sections 3, we call into question the surface level ratings attributed by the World Bank by cross-referencing with the on-ground realities of gender-based discrimination using data from Statistics Canada.
+The estimand we wish to replicate are the individual indicators used by WBL, but in a Canadian context that is backed by empirical research.
 The concluding section outlines the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.
@@
   geom_col() +
+  geom_text(aes(label = sprintf("%.2f", avg_index)), vjust = -0.5, color = "black") +  # Add this line for data labels
   labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
@@
 
-## Replication of Figure 1A: Avg. WBL Index in 2019
+## Replication of Figure 1B: Indicator variables averages 2019
 
-# average out the index across regions
-region_avg_index <- wb_data |> group_by(Region) |> filter(`WBL Report Year` == 2019) |>
-                    summarise(avg_index = mean(`WBL_index`, na.rm = TRUE), .groups = "drop")
+# isolate indicators 
+average_indicator <- wb_data |> 
+                        rename(
+                          Mobility = GR1_mobility, 
+                          Workplace = GR2_workplace, 
+                          Pay = GR3_pay, 
+                          Marriage = GR4_marriage, 
+                          Parenthood = GR5_parenthood, 
+                          Entrepreneurship = GR6_entrep, 
+                          Assets = GR7_assets, 
+                          Pension = GR8_pension
+                        ) |>
+                        filter(`WBL Report Year` == 2019) |>
+                        select(Mobility, 
+                               Workplace, 
+                               Pay, 
+                               Marriage, 
+                               Parenthood, 
+                               Entrepreneurship, 
+                               Assets, 
+                               Pension
+                               )
 
+#average them out
+avg_values <- colMeans(average_indicator)
+
+#put them in a table 
+new_table <- data.frame(Column_Names = names(avg_values), Average_Values = avg_values)
 # Plot in ggplot2
 
-  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
+  ggplot(data = new_table, mapping = aes(x = Column_Names, y = Average_Values)) +
   geom_col() +
-  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
+  labs(title="Indicator Averages, 2019", x="Indicator", y="WBL Indicator Score") +
   theme_minimal() +
@@
 \newpage
+Replication DOI Link: https://doi.org/10.48152/ssrp-qeab-5493
 

### 2024-04-23 outputs/paper/paper.qmd

@@
 date-format: long
-abstract: "Gender-based legal frameworks and workplace conditions are crucial considerations in the pursuit of global gender equality.
-This paper replicates and extends the data analysis conducted by Hyland et al.
-on the World Bank's 'Women, Business, and the Law' dataset.
-By cross-referencing the World Bank's indicators with Statistics Canada's data, we delve into the nuanced realities of gender-related legal frameworks and business regulations.
-The revelations underscore the need for a comprehensive and more practical approach in evaluating gender inclusivity within legal and business domains.
-"
+abstract: "Understanding gender-based differences in legal frameworks and workplace conditions is important in the pursuit of global gender equality.
+This paper replicates and extends Hyland, ASDFA, ASDFASF ().
+We examine gender-related legal frameworks and business regulations globally, and then compare the World Bank's indicators with those of Statistics Canada to focus on Canada.
+We find that nuances in country-level datasets are important for understanding the biases faced by women in different countries, which underscores the difficulty of constructing cross-country datasets."
 format: pdf
